---
title: "Final Presentation"
author: "Annika Kufrovich"
date: "January 24, 2019"
output: html_document
---

## Courses I took

* Intermediate R 
* Writing Functions in R 
* Writing Efficient R Code 
    + Looking at how to make sure the way you're coding is the fastest in can be 
    + When you've got a few hundred or a few thousand observations, this isn't a big deal 
    + When you have millions, you probably don't want to run code for 2 weeks 
* Parallel Programming 
    + An extension to the *fast* or *efficient* side of R coding 
    + Basically dividing up tasks between computer cores to reduce coding run time 
    + Can't be used on everything
* Working with geospatial data in R 
    + Really cool stuff, but sp objects are very confusing 
* Spatial Analysis in R with sf and raster
    + raster objects are gridded over large windows (elevation over an area could be in this)
    + sf objects are also like sp but much more intuitive, data frame with a list column to define shapes 
    + Data associated with sf and sp objects are meant for statistics of defined areas like counties, states, or nations (areal statistics) 
* Spatial Statitics 
    + Focuses on points or "events" within a "window" (a polygon from an sf or sp object)
    + Tests for distribution of events and density plots
    + Not limited to this, also goes through more things with areal statistics, space-time data, and GLMs
* Analyzing US Census Data 
    + Goes through how census data is structured and how to retrieve various kinds of census data 
* The last 4 courses **all** went through making pretty plots with ggplot2 



## Disclaimer and My data

Disclaimer:
I have been working with my data for a while, already had a pretty pecific idea of what I wanted to do with thi data during ISP, and have been coding in R for semesters

Data:
I have a big file of all registered voters in the state of Florida (This is public record because of Florida sunshine law). Each row is a registered voter and the variables are each a portion of the voter registration form. Variables I focus on are voter ID numbers, address, party affiliation, and active voter status. There are are roughly 30 more variable columns but we won't worry about those.



## Quick look at my matching function 

```{r, warning = FALSE, message = FALSE}
library(dplyr)
```

Test Dataframes

```{r, warning=FALSE, message=FALSE}
test1b <- data.frame(numbers = 1:8, letters = c("a", "b", "c", "d", "a", "b", "c", "d"))
test4b <- data.frame(numbers = 1:8, letters = c("a", "b", "c", "d", "a", "b", "c", "e"), 
                     lon = c(12, 31, 64, 38, 89, 73, 91, 21))
```

Function

```{r, warning=FALSE, message=FALSE}
matching_dfs_addcol <- function(newdf, olddf) {
  matching <- semi_join(olddf, newdf)
  matching_col <- matching %>%
    mutate(matched = TRUE)
  not_matching <- anti_join(newdf, olddf)
  not_matching_col <- not_matching %>%
    mutate(matched = FALSE)
  combined_df <- full_join(matching_col, not_matching_col)
  combined_df
}

new_func_test <- matching_dfs_addcol(test1b, test4b)
new_func_test
```


## The Pains of Geocoding


```{r, warning = FALSE, message = FALSE}
library(ggmap)
```

```{r, warning = FALSE, message = FALSE}
#loading in liberty county data
new_lib <- read.delim("C:/Users/TeenieTiny/Desktop/2nd Year/Data/Voter_Registration_20181113/20181113_VoterDetail/LIB_20181113.txt", header=FALSE)

#making an address column to use for geocoding
new_lib_address <- new_lib %>% 
  mutate(address = paste(V8, V10, "FL", sep = ", "))


#1st 1000 obs
nl_p1 <- new_lib_address %>% 
  slice(1:1000)
nl_geocode_p1 <- nl_p1 %>%
  mutate_geocode(address, source = "dsk")


#2nd 1000 obs
nl_p2 <- new_lib_address %>%
  slice(1001:2000)
nl_geocode_p2 <- nl_p2 %>%
  mutate_geocode(address, source = "dsk")


#3rd 1000 obs
nl_p3 <- new_lib_address %>%
  slice(2001:3000)
nl_geocode_p3 <- nl_p3 %>%
  mutate_geocode(address, source = "dsk")


#4th 1000 obs
nl_p4 <- new_lib_address %>%
  slice(3001:4000)
nl_geocode_p4 <- nl_p4 %>%
  mutate_geocode(address, source = "dsk")


#last few hundred obs
nl_p5 <- new_lib_address %>%
  slice(4001:4626)
nl_geocode_p5 <- nl_p5 %>%
  mutate_geocode(address, source = "dsk")

#joining
nl_join <- full_join(nl_geocode_p5, 
                     full_join(nl_geocode_p4, 
                               full_join(nl_geocode_p3, 
                                         full_join(nl_geocode_p1, nl_geocode_p2))))



#Then I realized I should have been using a function
geocoding <- function(df, col, seq) {
  sliced <- df %>% 
    slice(seq)
  geocoded <- sliced %>% 
    mutate_geocode(col, source = "dsk")
  geocoded
}

#ex <- geocoding(df = new_lib_address, col = address, seq = 1:3)
#ex$lon
#ex$lat

```


## Getting the shape of Liberty County

```{r, warning = FALSE, message = FALSE}
library(sp)
library(sf)
library(tigris)
library(spatstat)
library(maptools)
```


```{r, warning = FALSE, message = FALSE}

#retrieving sf object because it's easier to isolate liberty county this way
options(tigris_class = "sf")
fl_counties2 <- invisible(counties(state = "FL"))


#isolating liberty county
liberty <- fl_counties2 %>% 
  filter(NAME == "Liberty")
plot(liberty$geometry)


#retrieving sp object in case sf object doesn't behave
libdf <- as.data.frame(liberty$geometry[[1]][[1]][[1]])
colnames(libdf) <- c("x", "y")

libPoly <- Polygon(libdf, hole = FALSE)
libPolys <- Polygons(list(libPoly), ID = "liberty")
libSpPoly <- SpatialPolygons(list(libPolys))

liberty_window <- as.owin(W = libSpPoly)
plot(liberty_window)

nl_jxy <- nl_join %>% 
  mutate(x = lon, y = lat) %>% 
  mutate(party = ifelse(V24 == "DEM", "Democrat", 
                        ifelse(V24 == "REP", "Republican", "NPA/Other"))) %>%
  mutate(active = V29)

nl_jxy$party <- as.factor(nl_jxy$party)

nl_min <- nl_jxy %>% 
  select(x, y, party, active)

libppp <- ppp(nl_min$x, nl_min$y, window = liberty_window)
plot(libppp)

libppp_act <- ppp(nl_min$x, nl_min$y, window = liberty_window, marks = nl_min$active)
summary(libppp_act)

libppp_party <- ppp(nl_min$x, nl_min$y, window = liberty_window, marks = nl_min$party)
summary(libppp_party)
```



##Some pretty plots

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r, warning = FALSE, message = FALSE}
#nl_sites <- st_as_sf(nl_join, coords = c("lon", "lat"), 
#    crs = 4326, agr = "constant")
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

ggplot(data = world) + 
  geom_sf() + 
  geom_sf(data = fl_counties2, fill = NA, color = gray(0.5)) + 
  geom_point(data = nl_jxy, aes(x = lon, y = lat), color = "purple", alpha = 0.4) + 
  coord_sf(xlim = c(-85.5, -84), ylim = c(29.5, 31)) + 
  theme_light() + 
  labs(x = "Longitude", y = "Latitude", title = "Registered Voters", 
       subtitle = "Liberty County, FL")


ggplot(data = world) + 
  geom_sf() + 
  geom_sf(data = fl_counties2, fill = NA, color = gray(0.5)) + 
  geom_point(data = nl_jxy, aes(x = lon, y = lat, col = active), shape = "o", alpha = 0.4) + 
  coord_sf(xlim = c(-85.5, -84), ylim = c(29.5, 31)) + 
  scale_color_manual(values = c("coral", "black")) + 
  theme_light() + 
  labs(x = "Longitude", y = "Latitude", title = "Active vs Inactive Voters", 
       subtitle = "Liberty County, FL", colour = "Status")


ggplot(data = world) + 
  geom_sf() + 
  geom_sf(data = fl_counties2, fill = NA, color = gray(0.5)) + 
  geom_point(data = nl_jxy, aes(x = lon, y = lat, col = party), shape = "o", alpha = 0.4) + 
  coord_sf(xlim = c(-85.5, -84), ylim = c(29.5, 31)) + 
  scale_color_manual(values = c("blue", "green", "red")) + 
  theme_light() + 
  labs(x = "Longitude", y = "Latitude", title = "Party Affiliation of Voters", 
       subtitle = "Liberty County, FL", colour = "Party")

ggplot(data = world) + 
  geom_sf() + 
  geom_sf(data = fl_counties2, fill = NA, color = gray(0.5)) + 
  geom_point(data = nl_jxy, aes(x = lon, y = lat, col = party), shape = "o", alpha = 0.4) + 
  coord_sf(xlim = c(-85.5, -84), ylim = c(29.5, 31)) + 
  scale_color_manual(values = c("light blue", "green", "red")) + 
  theme_light() + 
  labs(x = "Longitude", y = "Latitude", title = "Party Affiliation of Voters", 
       subtitle = "Liberty County, FL", colour = "Party")


```


## Density plots

```{r, warning = FALSE, message = FALSE}
plot(libppp)
gen_dens <- plot(density(libppp), main = "Density of registered voters in Liberty County, FL")
invisible(gen_dens)

active_splits <- split(libppp_act) 
bas_act <- plot(active_splits, main = "Active and Inactive Voters")
invisible(bas_act)

dens_act <- density(active_splits)
plot(dens_act, main =  "Active and Inactive Voters")

frac_dens_act <- dens_act[[2]] / (dens_act[[1]] + dens_act[[2]])
plot(frac_dens_act, main = "Density for proportion of Inactive voters")

```

## Future Research
* Figure out why so many points were outside liberty county and somw were not gecoded 
* Expand to other counties in Florida, Florida as a whole
* Use data for various GLM's
* Work on more accurate and more efficient geocoding