---
title: "FunctionNotes"
author: "Annika Kufrovich"
date: "January 6, 2019"
output: html_document
---

##Notes

* Reviewed when to write a function, generally when code is repeated more than twice and code is bulky 
* Functions should have minimal duplication, mainly for clarity 
* Good functions are correct and understandable 
* You can make arguments for operations within your function 
* Using purrr 
    + map functions work similarly to the apply functions and each has a different return type 
    + `map()` returns a list 
    + `map_dbl()` returns a double vector 
    + `map_lgl()` returns a logical vector 
    + `map_int()` returns an integer vector 
    + `map_chr()` returns a character vector 
    + `~` followed by a formula can replace calling a defined function or an anonymous function, `.` would be used as a place holder for an element of `.x` or vector argument 
    + you can also shorthand subsetting in lists 
    + sometimes the map functions fail and won't show successes or failures, you can see sucesses and failures if you wrap these functions in the purrr function `safely()` 
    + `possibly()` is similar to `safely()` in making map functions run, but intead of making two lists, you define a replacement for iterations with errors 
    + `quietly()` ets all extra output produced by a function 
    + `map2()` allows for a second object/argument to be iterated over 
    + `pmap()` allows for any number of arguments to be iterated over, has a list as its first argument `.l` 
    + `invoke_map()` is used for iterating over functions and , order of aruments is reversed now taking a list of functions first for `.f`, and an argument for `.x` second 
    + `walk()` has same arguments as map functions but is used for side effects, like saving different ggplots to pdfs
    + lists are a lot 
* Functions hould have clear error mesages when they fail 
* 3 main problematic functions: 
    + Type-unstable functions, ie functions that don't have a set type of output 
    + Non-standard evaluation, when things have ambiguity 
    + Hidden arguments 
* Using `stop()` in a function allows you to have infromative error messaes for different conditions (when placed in an if statement)
* This course felt very theoretical and less practical, would have prefered work with functions that had more to do with dataframes 


##Practice/Application


```{r, include=FALSE}
library(purrr)
library(compare)
library(dplyr)
```


Comparing test dataframes 

```{r, warning=FALSE, message=FALSE}
test1 <- data.frame(numbers = 1:8, letters = c("a", "b", "c", "d", "a", "b", "c", "d"))
test2 <- rev(test1)
comparison <- compare::compare(test1, test2, allowAll = TRUE)
comparison$tM
```

```{r, warning=FALSE, message=FALSE}
test1a <- data.frame(numbers = 1:8, letters = c("a", "b", "c", "d", "a", "b", "c", "d"))
test2a <- data.frame(numbers = 1:7, letters = c("a", "b", "c", "d", "a", "b", "c"))
comparison_a <- compare::compare(test1a, test2a, allowAll = TRUE)
comparison_a$tM
```


while the compare function is cool, I couldn't figure out how to use its output to make a new column. It didn't have an attribute with a data frame of the unmatched rows from test1a.  


```{r, warning=FALSE, message=FALSE}
test1b <- data.frame(numbers = 1:8, letters = c("a", "b", "c", "d", "a", "b", "c", "d"))
test2b <- data.frame(numbers = 1:8, letters = c("a", "b", "c", "d", "a", "b", "c", "e"))
comparison_b <- semi_join(test1b, test2b)
comparisonb_match <- comparison_b %>% 
  mutate(matched = TRUE)
comparisonb2 <- anti_join(test1b, test2b)
comparisonb2_unmatch <- comparisonb2 %>% 
  mutate(matched = FALSE)
comparisonb2_unmatch
```


retest with reversed test2b

```{r, warning=FALSE, message=FALSE}
test3b <- rev(test2b)
comparison_b_re <- semi_join(test1b, test3b)
comparisonb_match_re <- comparison_b_re %>% 
  mutate(matched = TRUE)
comparisonb2_re <- anti_join(test1b, test3b)
comparisonb2_unmatch_re <- comparisonb2_re %>% 
  mutate(matched = FALSE)


matching_status <- full_join(comparisonb_match_re, comparisonb2_unmatch_re)
matching_status
```

It works! now to make a function

```{r, warning=FALSE, message=FALSE}
matching_dfs <- function(newdf, olddf) {
  matching <- semi_join(newdf, olddf)
  matching_col <- matching %>%
    mutate(matched = TRUE)
  not_matching <- anti_join(newdf, olddf)
  not_matching_col <- not_matching %>%
    mutate(matched = FALSE)
  combined_df <- full_join(matching_col, not_matching_col)
  combined_df
}

func_test <- matching_dfs(test1b, test3b)
func_test
```

Testing with "olddf" argument having an extra column

```{r, warning=FALSE, message=FALSE}
test4b <- data.frame(numbers = 1:8, letters = c("a", "b", "c", "d", "a", "b", "c", "e"), lon = c(12, 31, 64, 38, 89, 73, 91, 21))

func_test2 <- matching_dfs(test1b, test4b)
func_test2
```

Now to figure out how to take extra column values in matched rows and add them to the new df. Unmatched values should have NA in this column

```{r, warning=FALSE, message=FALSE}
matching_dfs_addcol <- function(newdf, olddf) {
  matching <- semi_join(olddf, newdf)
  matching_col <- matching %>%
    mutate(matched = TRUE)
  not_matching <- anti_join(newdf, olddf)
  not_matching_col <- not_matching %>%
    mutate(matched = FALSE)
  combined_df <- full_join(matching_col, not_matching_col)
  combined_df
}

new_func_test <- matching_dfs_addcol(test1b, test4b)
new_func_test
```




##Goals for Tomorrow

* Test my function with a small county from the voter file tomorrow
* Do first 2-3 chapters of Efficient R Coding course on Data camp
* start trying to apply ideas to the functions I have created. Maybe look for base r alternatives since 2nd chapter is focused on fast base r code.
* Dicuss with Professor Crow New ISP plans, maybe ask for resources on PostgreSQL and PostGIS